CTest Integration of Sikuli Automated GUI Testing

### 2011-06-16 Documentation/Article.tex
edit

Figure 1 shows the Sikuli-IDE for version 0.10.2.  On the left-hand side menu is a list
of commonly used Sikuli commands like  \emph{find}(),  \emph{click}(), and  \emph{type}().  The picture of
the camera within the parentheses indicates the screenshot input variable.  Clicking
on any of these menu functions will automatically set up the Sikuli screen shot mode, whereby
the Sikuli-IDE disappears and what is on the screen at that moment is available to capture 
by selecting a rectangular region of interest.  The function will then appear in the scripting domain
with the snap shot in parentheses as shown in Figure 1.  The user may also type in the function
and invoke the screen shot mode by clicking on the camera in the top menu or by using a
keyboard shortcut. The simple Sikuli script example in Figure 1 will open gedit Text Editor, type a message,
and save the file. (see Figure~\ref{fig:SimpleExample}).

@@

@@

The newest versions of Sikuli come equipped with a Unit Testing feature based
on JUnit~\cite{}. Unit Testing can be called using the IDE or by command
line and enables the user to write multiple functions in a script which are
run individually. At the end of the run, the Sikuli output will tell the user
how many tests ran successfully and which ones failed and where.\\

Here we connect this Unit Testing feature to CTest, this allows us to report
results on public dashboard using CDash.

The Sikuli script in Figure 2 shows a fragment of a unit test for the GoFigure2 software.  The unit being tested here
is the Toolbar menu which consists of different modes used to analyze data set plots.
In total there are four modes which are represented in the first line as the vector Toolbar and for each
of these modes we want to test their expected mouse functions.  For instance, if we are in the zoom Toolbar mode
then we want the data set plot to zoom in when the user scrolls the mouse wheel forward.  Likewise,
if we are in pan mode then we want the plot to move to the left if the user holds down the left mouse button and moves
the mouse to the left.  Also importantly, we want to see that these two mouse actions are unique for their respective Toolbar modes;
the test will include an assertion that scrolling forward will zoom in when using zoom mode and will do nothing when using pan mode.\\

In addition to representing the modes and actions
as the vectors Toolbar and Mouse, respectively, we've included the different data set plots in the vector ViewRegion.  
These images represent the  \emph{xy-},  \emph{xz-},  \emph{yz-}, and  \emph{xyz-}views of zebrafish microscopy data taken in the Megason Lab.  After capturing
these images we can then use them to locate their initial pixel coordinates within the GoFigure2 GUI using the \emph{find}() function.  These points become 
our reference positions to calculate directional movement.  For example, when testing the pan mode, we mark the initial position of each image,
use the \emph{mouseMove}() function to a chosen location, and then use \emph{find}() to obtain the new coordinates.  We can then assert if the image panned as it should have.
Similarly for zoom, we can first find the area of the image and then find the area of the zoomed in image.  To test the result when the user clicks 
in a certain mode, we've made use of Sikuli's \emph{assert exists}() function to ascertain the existence of our
desired results   (see Figure~\ref{fig:Gofigure2Example}).\\

\begin{figure}[htp]
 \centering
 \includegraphics[width=0.99\textwidth]{Images/Gofigure2Example.png}
 % Gofigure2Example.png: 667x826 pixel, 72dpi, 23.53x29.14 cm, bb=0 0 667 826
 \caption{Sikuli Script for Gofigure2 GUI Testing}
 \label{fig:Gofigure2Example}
\end{figure}
% ------------------------------------------------------------------------
\section{Conclusion}

>>>>

Figure~\ref{fig:SimpleExample} shows the Sikuli-IDE for version 0.10.2. On the
left-hand side menu is a list of commonly used Sikuli commands
like \emph{find}(), \emph{click}(), and \emph{type}(). The picture of
the camera within the parentheses indicates the screenshot input variable.
Clicking on any of these menu functions will automatically set up the Sikuli
screen shot mode, whereby the Sikuli-IDE disappears and what is on the screen at
that moment is available to capture by selecting a rectangular region of
interest. The function will then appear in the scripting domain with the snap
shot in parentheses as shown in Figure~\ref{fig:SimpleExample}. The user may
also type in the function and invoke the screen shot mode by clicking on the
camera in the top menu or by using a keyboard shortcut. The simple Sikuli script
example in Figure~\ref{fig:SimpleExample} will open gedit Text Editor, type a
message, and save the file.

@@

The Sikuli script in Figure~\ref{fig:Gofigure2Example} shows a fragment of a
unit test for the GoFigure2 software~\cite{GoFigure2:Website}. The unit being
tested here is the Toolbar menu which consists of different modes used to
analyze data set plots. In total there are four modes which are represented in
the first line as the vector Toolbar and for each of these modes we want to test
their expected mouse functions. For instance, if we are in the zoom Toolbar
mode then we want the data set plot to zoom in when the user scrolls the mouse
wheel forward. Likewise, if we are in pan mode then we want the plot to move to
the left if the user holds down the left mouse button and moves the mouse to the
left.  Also importantly, we want to see that these two mouse actions are unique
for their respective Toolbar modes; the test will include an assertion that
scrolling forward will zoom in when using zoom mode and will do nothing when
using pan mode.\\

In addition to representing the modes and actions as the vectors Toolbar and
Mouse, respectively, we have included the different data set plots in the
vector ViewRegion. These images represent the \emph{xy-}, \emph{xz-},
\emph{yz-}, and \emph{xyz-}views of zebrafish microscopy data taken in the
Megason Lab. After capturing these images we can then use them to locate their
initial pixel coordinates within the GoFigure2 GUI using the \emph{find}()
function. These points become our reference positions to calculate directional
movement.  For example, when testing the pan mode, we mark the initial position
of each image, use the \emph{mouseMove}() function to a chosen location, and
then use \emph{find}() to obtain the new coordinates. We can then assert if the
image panned as it should have. Similarly for zoom, we can first find the area
of the image and then find the area of the zoomed in image.  To test the result
when the user clicks in a certain mode, we've made use of Sikuli's \emph{assert
exists}() function to ascertain the existence of our desired results (see
Figure~\ref{fig:Gofigure2Example}).\\

\begin{figure}[htp]
 \centering
 \includegraphics[width=0.99\textwidth]{Images/Gofigure2Example.png}
 % Gofigure2Example.png: 667x826 pixel, 72dpi, 23.53x29.14 cm, bb=0 0 667 826
 \caption{Sikuli Script for Gofigure2 GUI Testing}
 \label{fig:Gofigure2Example}
\end{figure}

@@

The newest versions of Sikuli come equipped with a Unit Testing feature based
on JUnit~\cite{JUnit:Website}. Unit Testing can be called using the IDE or by
command line and enables the user to write multiple functions in a script which
are run individually. At the end of the run, the Sikuli output will tell the
user how many tests ran successfully and which ones failed and where.\\

Here we connect this Unit Testing feature to CTest, this allows us to report
results on public dashboard using CDash. To this end, we developed cmake
functions to discover where Sikuli is installed and special function to add
sikuli test in the list of test for CTest.

Here is a typical CMakeLists.txt file, to be able to use such functionalities:

\begin{verbatim}
# first let's include CTest and enable testing

include( CTest )
enable_testing()

# Then let's
#   - figure out where Sikuli is installed
#   - define function to add Sikuli tests

find_package( Sikuli )

# here we add a sikuli test
add_sikuli_test( ThunderbirdSikuliTest test.sikuli )
\end{verbatim}

Users can either maintain an image library, either keep images in the
corresponding sikuli script directory (default). If one use an image library,
the corresponding path must be set during CMake time in
the SIKULI\_IMAGE\_LIBRARY\_DIR variable.

% ------------------------------------------------------------------------
\section{Conclusion}
